---
title: "Astronaut Database - Tidy Tuesday Wk 29 2020"
author: "Joe Roberts"
date: "15/01/2021"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

## Load packages
library(tidyverse)
library(rnaturalearth)
library(rnaturalearthdata)
library(rgeos)
library(sf)
library(rworldmap)
library(patchwork)
library(Cairo)
library(extrafont)
library(plotly)
```

## Data background 
This database contains publically available information about all astronauts who participated in space missions before 15 January 2020 collected from NASA and Roscosmos. The provided information includes full astronaut name, sex, date of birth, nationality, military status, a title and year of a selection program, and information about each mission completed by a particular astronaut such as a year, ascend and descend shuttle names, mission and extravehicular activity (EVAs) durations. More detail is outlined in the associated peer-reviewed publication by [Corlett, Stavnichuk and Komarova (2020)](https://doi.org/10.1016/j.lssr.2020.06.003).

## Import and inspect data
```{r cars}
# Download data from Tidy Tuesday GitHub repository
astronaut_data <- readr::read_csv('https://raw.githubusercontent.com/rfordatascience/tidytuesday/master/data/2020/2020-07-14/astronauts.csv')

# Inspect data structure
# str(astronaut_data)

# Data summary 
# summary(astronaut_data)
```

## Exploratory data analysis
Which countries have produced the largest number of astronauts? This is likely to be the USA and Russia, but let's confirm this:

```{r}
# Extract astronaut's country of origin from data and calculate the count 
countries = data.frame(table(astronaut_data$nationality))

# Define 'better' colnames
colnames(countries) = c("ctry_name", "nr_of_astros")

# Plot data
ggplot(data = countries, aes(x = reorder(ctry_name, nr_of_astros), y = nr_of_astros)) +
  geom_col() +
  coord_flip() +
  theme_minimal()
```

From this plot it is obvious that there are some issues with the country names. These need to be fixed:

```{r}
# Change the country expression
countries$ctry_name = gsub("Czechoslovakia", "Slovakia", countries$ctry_name)
countries$ctry_name = gsub("Hungry", "Hungary", countries$ctry_name)
countries$ctry_name = gsub("Malysia", "Malaysia", countries$ctry_name)
countries$ctry_name = gsub("Netherland", "Netherlands", countries$ctry_name)
countries$ctry_name = gsub("U.S.S.R/Russia", "Russia", countries$ctry_name)
countries$ctry_name = gsub("U.S.S.R/Ukraine", "Ukraine", countries$ctry_name)
countries$ctry_name = gsub("Korea", "South Korea", countries$ctry_name)
countries$ctry_name = gsub("Republic of South Africa", "South Africa", countries$ctry_name)
countries$ctry_name = gsub("U.K.", "United Kingdom", countries$ctry_name)
countries$ctry_name = gsub("U.S.", "United States of America", countries$ctry_name)
countries$ctry_name = gsub("UAE", "United Arab Emirates", countries$ctry_name)

# Filter to remove double citizenship
countries = countries %>%
  filter(ctry_name != "United Kingdom/United States of America")

# Replot data to ensure these changes have been implemented
ggplot(data = countries, aes(x = reorder(ctry_name, nr_of_astros), y = nr_of_astros)) +
  geom_col() +
  coord_flip() +
  theme_minimal()
```

## Visualising astronaut origin using a map
This data could be visualised with more impact on a map!

```{r}
# Data prep----

# Get worldmap
world_map <- getMap(resolution="high")

# Get centroids per country polygon
centroids <- gCentroid(world_map, byid=TRUE)

# Define df with centroids including x and y coordinate
centroid_df <- data.frame(centroids)

d <- cbind(rownames(centroid_df), data.frame(centroid_df, row.names=NULL))
colnames(d) = c("ctry_name", "x", "y")

# Join df on countries
countries_geo <- left_join(countries, d,
                           by = "ctry_name")

# Convert df to sf object
countries_geo_sf <- sf::st_as_sf(countries_geo, coords = c("x","y")) 

# Change crs
st_crs(countries_geo_sf) = 4326


# data visualisation----

# Download country polygons
world <- ne_countries(scale = "medium", returnclass = "sf")

map <- ggplot() +
  # Add country polygons
  geom_sf(data = world, fill = "#FFFFFF", color = "#2A363B", size = .2) +
  
  # Add point centroid locations
  geom_sf(data = countries_geo_sf, color = "black", size = 3, alpha = .5) +
  geom_sf(data = countries_geo_sf, color = "red", size = 1, alpha = .75) +

  # Add title and subtitle
  labs(subtitle = "Astronaut's Countries of Origin\n",
       title = "\nWHERE IN THE WORLD?") +
  
  # Get globe projection, set to northpole and europe
  coord_sf(crs = "+proj=laea +lat_0=52 +lon_0=10 +x_0=4321000 +y_0=3210000 +ellps=GRS80 +units=m +no_defs") +
  
  # Theming
  theme_minimal() +
  theme(plot.background = element_rect(fill = "#000000",
                                       color ="#000000"),
        panel.grid = element_line(color = "#FFFFFF"),
        plot.title = element_text(colour = "#FFFFFF",
                                  hjust = .5,
                                  family = "Bell MT",
                                  size = 20),
        plot.subtitle = element_text(colour ="#FFFFFF",
                                     hjust = .5,
                                     family = "Verdana"),
        panel.grid.major = element_line(size = .1))

#save plot
ggsave(map, filename = "globe.png", width = 14.8, height = 19, units = "cm", type = "cairo-png", dpi = 300)
```

## How old are astronauts on their first mission?

```{r}
library(tidyverse)
library(extrafont)
library(ggdark)
library(devtools)
library(viridis)

age_astronauts <- astronaut_data %>% 
  mutate(age_at_mission = year_of_mission - year_of_birth) %>% 
  select(name, nationality, sex, age_at_mission, hours_mission, mission_number, year_of_mission, occupation)

ggplot(age_astronauts, aes(x = year_of_mission, y = age_at_mission, colour = mission_number)) + 
  geom_jitter(aes(alpha = 0.7)) +
  theme_minimal() +
  labs(title = "How old are astronauts when they go into space?") +
  scale_y_continuous(name = "Age at mission launch") +
  scale_color_viridis(option = "plasma") +
  theme(plot.background = element_rect(fill = "#000000"),
        panel.background = element_blank(),
        panel.grid.major = element_line(color = "grey30", size = 0.2),
        panel.grid.minor = element_line(color = "grey30", size = 0.2),
        legend.background = element_blank(),
        axis.ticks = element_blank(),
        axis.text = element_text(color = "#C7C4C4", size = 20),
        legend.key = element_blank(),
        legend.title = element_text(color = "#C7C4C4", size = 16),
        legend.text = element_text(color = "#C7C4C4", size = 16),
        axis.title.x = element_blank(),
        axis.title.y = element_text(color = "#C7C4C4", size=20, margin = margin(0,30,0,0)),
        plot.title = element_text(color="#C7C4C4", size=20, face="bold",margin = margin(0,0,10,0)),
        plot.subtitle = element_text(color = "#C7C4C4", size=12, face="bold.italic", margin = margin(0,0,30,0)),
        plot.caption = element_text( color = "grey30", size=8),
        plot.margin=unit(c(1,1,1.5,1.2),"cm")) +
  scale_alpha(guide = 'none') +
  guides(color=guide_legend(title="Mission Number")) +
  annotate("text", x = 1985, y = 70,
    label = "John Glenn (77)\n is the oldest person to\ntravel to space",
    colour = "white",
    vjust = 1,
    size = 5
  ) + 
  annotate(
    "curve",
    xend = 1997,
    yend = 77,
    x = 1990,
    y = 70,
    curvature = -.2,
    arrow = arrow(type = "closed", length = unit(0.40, "lines")),
    colour = "white"
  ) + annotate("text", x = 1980, y = 27,
           label = "Guerman Titov and\n Valentina Tereshkova (26)\n are the youngest people to travel to space",
           colour = "white",
           vjust = 1,
           size = 5
  ) +
  annotate(
    "curve",
    xend = 1961,
    yend = 25,
    x = 1970,
    y = 25,
    curvature = -.2,
    arrow = arrow(type = "closed", length = unit(0.40, "lines")),
    colour = "#C7C4C4"
  ) +
  annotate(
    "curve",
    xend = 1963,
    yend = 25,
    x = 1970,
    y = 25,
    curvature = -.2,
    arrow = arrow(type = "closed", length = unit(0.40, "lines")),
    colour = "#C7C4C4"
  ) +
  geom_smooth(aes(group = 1), method = "lm", formula = y ~ x, se = FALSE, colour = "grey", alpha = 0.7)

ggsave("age_astronauts.png", dpi = 300, width = 11, height = 11) 
```

